Quizzes
========================================================

## Week 1 Quiz 

### Introduction

This first quiz will check your ability to execute basic operations on objects in R and to understand some basic concepts. For questions 11-20 you will need to load a dataset into R and do some basic manipulations in order to answer the questions on the quiz.

You may want to print a copy of the quiz questions to look at as you work on the assignment. It is recommended that you save your answers as you go in the event that a technical problem should occur with your network connection or computer. Ultimately, you must submit the quiz online to get credit!

### Data

The zip file containing the data for questions 11-20 in this Quiz can be downloaded here:

1. Week 1 Quiz Data

For this assignment you will need to unzip this file in your working directory.

**Question 1**

The R language is a dialect of which of the following programming languages?

**Answer**

S

**Question 2**

The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition?

**Answer**

The freedom to restrict access to the source code for the software.

**Question 3**

In R the following are all atomic data types EXCEPT 

**Answer**

R has five *atomic* classes:

- Character
- Numeric (real numbers)
- Integer
- complex
- Logical (True/False)

Therefore, Matrix is not an atomic data type.

**Question 4**

If I execute the expression x <- 4 in R, what is the class of the object *x* as determined by the `class()' function?

**Answer**

```{r class type}

x = 4
class(x)

```
Numeric

**Question 5**

What is the class of the object defined by x <- c(4, TRUE)?


```{r class type}

x = c(4, TRUE)
x <- c(4, "a", TRUE)
class(x)

x <- c(1,3, 5)
y <- c(3, 2, 10)
cbind(x,y)
x <- list(2, "a", "b", TRUE)
x[[2]]
```

Numeric

**Question 6**

If I have two vectors x <- c(1,3, 5) and y <- c(3, 2, 10), what is produced by the expression rbind(x, y)?

**Answer**

```{r class type}

x = c(1, 3,5)
y = c(3,2,10)
rbind(x,y)
```
A matrix with two rows and three columns

**Question 7**

A key property of vectors in R is that:

**Answer**

elements of a vector all must be of the same class

**Question 8**

Suppose I have a list defined as x <- list(2, "a", "b", TRUE). What does x[[1]] give me?

**Answer**

```{r list example}
x <- list(2, "a", "b", TRUE)
class(x[[1]])
class(x[[2]])
x[1]
```

**Question 9**

Suppose I have a vector x <- 1:4 and y <- 2:3. What is produced by the expression x + y?

```{r addition of two vectors }
x <- 1:4
y <- 2:3
x
y
x+y
```
an numeric vector with the values 3, 5, 5, 7

**Question 10**

Suppose I have a vector x <- c(17, 14, 4, 5, 13, 12, 10) and I want to set all elements of this vector that are greater than 10 to be equal to 4. What R code achieves this?

```{r changing vector values}
x <- c(17, 14, 4, 5, 13, 12, 10)
x
x[x>10] = 4
x
```
x[x>10] = 4

**Question 11**

In the dataset provided for this Quiz, what are the column names of the dataset?

```{r read the dataset}

quiz1 = read.csv("hw1_data.csv")
str(quiz1)
```
Ozone, Solar.R, Wind, Temp, Month, Day

**Question 12**

Extract the first 2 rows of the data frame and print them to the console. What does the output look like?

```{r data extraction}
head(quiz1)
quiz1[1:2,]
```

**Question 13**

How many observations (i.e. rows) are in this data frame?

**Answer**

153

**Question 14**

Extract the last 2 rows of the data frame and print them to the console. What does the output look like?

```{r data extraction 2}
tail(quiz1)
quiz1[152:153,]
```

**Question 15**

What is the value of Ozone in the 47th row?

**Answer**

```{r row data extraction}
tail(quiz1)
quiz1$Ozone[47]
```
21

**Question 16**

How many missing values are in the Ozone column of this data frame?

```{r missing values}
summary(quiz1$Ozone)

```
37

**Question 17**

What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.

**Answer**

42.13

**Question 18**

Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?

```{r subset data}

quiz1subset = subset(quiz1, quiz1$Ozone > 31 & quiz1$Temp > 90)
summary(quiz1subset$Solar.R)
```

**Question 19**

What is the mean of "Temp" when "Month" is equal to 6? 

```{r mean}

tapply(quiz1$Temp, quiz1$Month == 6, mean)
```

**Question 20**

What was the maximum ozone value in the month of May (i.e. Month = 5)?

```{r max value}
quiz1subset = subset(quiz1, quiz1$Month == 5)
max(quiz1subset$Ozone, na.rm=TRUE)
```