Quiz 4
=============

## Question 1
What is produced at the end of this snippet of R code?

```{r Q1}
set.seed(1)
rpois(5, 2)

```
**ANS**: A vector with the numbers 1, 1, 2, 4, 1, because the *set.seed()* function is used, *rpois()* will always output the same vector in this code.

## Question 2
What R function can be used to generate standard Normal random variables?

rnorm
pnorm
dnorm
qnorm

```{r Q2}
rnorm
dpois
```
**ANS**:rnorm, Functions beginning with the `r' prefix are used to simulate random variates. Standard probability distributions in R have a set of four functions that can be used to simulate variates, evaluate the density, evaluate the cumulative density, and evaluate the quantile function.

## Question 3

When simulating data, why is using the set.seed() function important?

1. It can be used to specify which random number generating algorithm R should use.
2. It ensures that the sequence of random numbers is truly random.
3. It can be used to generate non-uniform random numbers.
4. It ensures that the random numbers generated are within specified boundaries. 

**ANS**: 1, It ensures that the sequence of random numbers is reproducible.


## Question 4

Which function can be used to evaluate the inverse cumulative distribution function for the Poisson distribution?
1. dpois
2. qpois
3. rpois
4. ppois

**ANS**: 2, Probability distribution functions beginning with the `q' prefix are used to evaluate the cumulative distribution function.

## Question 5

What does the following code do?

```{r Q5}
set.seed(10)
x <- rbinom(10, 10, 0.5)
e <- rnorm(10, 0, 20)
y <- 0.5 + 2 * x + e

```

1. Generate random exponentially distributed data
2. Generate data from a Poisson generalized linear model
3. Generate uniformly distributed random data
4. Generate data from a Normal linear model 

**ANS**: 4

## Question 6

What R function can be used to generate Binomial random variables?

pbinom
dbinom
qbinom
rbinom

**ANS**: rbinom

## Question 7

What aspect of the R runtime does the profiler keep track of when an R expression is evaluated?

the working directory
the global environment
the function call stack
the package search list

**ANS**: the function call stack

## Question 8

Consider the following R code:

```{r Q8}
library(datasets)
Rprof()
fit <- lm(y ~ x1 + x2)
Rprof(NULL)
```
(Assume that y, x1, and x2 are present in the workspace.) Without running the code, what percentage of the run time is spent in the 'lm' function, based on the 'by.total' method of normalization shown in 'summaryRprof()'?

It is not possible to tell
100%
23%
50% 

**ANS**: 100%, When using *by.total* normalization, the top-level function (in this case, `lm()') always takes 100% of the time.

## Question 9

When using 'system.time()', what is the user time?

It is the time spent by the CPU evaluating an expression
It is the time spent by the CPU waiting for other tasks to finish
It is the "wall-clock" time it takes to evaluate an expression
It is a measure of network latency

**ANS**: It is the time spent by the CPU evaluating an expression

## Question 10

If a computer has more than one available processor and R is able to take advantage of that, then which of the following is true when using 'system.time()'?

user time is 0
elapsed time is 0
user time is always smaller than elapsed time
elapsed time may be smaller than user time 

**ANS**: elapsed time may be smaller than user time 
