Programming Assignment 1: Air Pollution: Instructions
========================================================

## Introduction

For this first programming assignment you will write three functions that are meant to interact with dataset that accompanies this assignment. The dataset is contained in a zip file **specdata.zip** that you can download from the Coursera web site. 

## Data

The zip file containing the data can be downloaded here: 

specdata.zip [2.4MB] 

The zip file contains 332 comma-separated-value (CSV) files containing pollution monitoring data for fine particulate matter (PM) air pollution at 332 locations in the United States. Each file contains data from a single monitor and the ID number for each monitor is contained in the file name. For example, data for monitor 200 is contained in the file "200.csv". Each file contains three variables:

* Date: the date of the observation in YYYY-MM-DD format (year-month-day)
* sulfate: the level of sulfate PM in the air on that date (measured in micrograms per cubic meter)
* nitrate: the level of nitrate PM in the air on that date (measured in micrograms per cubic meter)

For this programming assignment you will need to unzip this file and create the directory 'specdata'. Once you have unzipped the zip file, do not make any modifications to the files in the 'specdata' directory. In each file you'll notice that there are many days where either sulfate or nitrate (or both) are missing (coded as NA). This is common with air pollution monitoring data in the United States. 

## Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA. A prototype of the function is as follows:

```{r prototype of Part 1}
pollutantmean <- function(directory, pollutant, id = 1:332) {
        ## 'directory' is a character vector of length 1 indicating
        ## the location of the CSV files
        
        ## 'pollutant' is a character vector of length 1 indicating
        ## the name of the pollutant for which we will calculate the
        ## mean; either "sulfate" or "nitrate".
        
        ## 'id' is an integer vector indicating the monitor ID numbers
        ## to be used
        
        ## Return the mean of the pollutant across all monitors list
        ## in the 'id' vector (ignoring NA values)
        }

```

The output should look like:

```{r part 1 output}

source("pollutantmean.R")
pollutantmean("specdata", "sulfate", 1:10)

## [1] 4.064

pollutantmean("specdata", "nitrate", 70:72)

## [1] 1.706

pollutantmean("specdata", "nitrate", 23)

## [1] 1.281


```

```{r simple function assignment part 1}

pollutantmean <- function(directory, pollutant, id = 1:332) {
        
        #set the path
       path = "C:/Users/Fahad/Documents/R Projects/Coursera/R-Programming-Coursera/Data/"
        path = c(path,directory)
        path = paste(path,collapse="")
        #get the file List in that directory
        fileList = list.files(path)
        
        #stripping off the extension part and extract the file names 
        file.names = as.numeric(sub("\\.csv$","",fileList))
        
        #select files to be imported based on the user input or default
        selected.files = fileList[match(id,file.names)]
        
        #import data
        Data = lapply(file.path(path,selected.files),read.csv)
        
        #convert into data frame
        Data = do.call(rbind.data.frame,Data)
        
        #calculate mean
        mean(Data[,pollutant],na.rm=TRUE)
        
        }
        
pollutantmean("specdata","nitrate",70:72)
```

## Part 2

Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows

```{r assign part2 prototype}
complete <- function(directory, id = 1:332) {
        ## 'directory' is a character vector of length 1 indicating
        ## the location of the CSV files

        ## 'id' is an integer vector indicating the monitor ID numbers
        ## to be used
        
        ## Return a data frame of the form:
        ## id nobs
        ## 1  117
        ## 2  1041
        ## ...
        ## where 'id' is the monitor ID number and 'nobs' is the
        ## number of complete cases
}
```

The output should look like:

```{r assign p2 output}
source("complete.R")
complete("specdata", 1)

##   id nobs
## 1  1  117

complete("specdata", c(2, 4, 8, 10, 12))

##   id nobs
## 1  2 1041
## 2  4  474
## 3  8  192
## 4 10  148
## 5 12   96

complete("specdata", 30:25)

##   id nobs
## 1 30  932
## 2 29  711
## 3 28  475
## 4 27  338
## 5 26  586
## 6 25  463

complete("specdata", 3)

##   id nobs
## 1  3  243
```

he function that you write should be able to match this output. Please save your code to a file named **complete.R**. To run the submit script for this part, make sure your working directory has the file **complete.R** in it. 

```{r assign p2 solution}

complete <- function(directory, id = 1:332) {
        
        #set the path
        path = "C:/Users/607518069/Documents/R Projects/Coursera/R-Programming-Coursera/Data/"
        path = c(path,directory)
        path = paste(path,collapse="")
        #get the file List in that directory
        fileList = list.files(path)
        
        #extract the file names and store as numeric for comparison
        file.names = as.numeric(sub("\\.csv$","",fileList))
        
        #select files to be imported based on the user input or default
        selected.files = fileList[match(id,file.names)]
        
        #import data
        Data = lapply(file.path(path,selected.files),read.csv)
        
        #get the complete cases
        completeCases = lapply(Data, function(Data) Data[complete.cases(Data),])
        
        #create a dataframe to return
        res = data.frame()
        
        #get the number of complete cases and put it in the result dataframe
        for(i in 1:length(completeCases)){
                res <- rbind(res, nob=nrow(completeCases[[i]]))
        }
        
        #add ID column
        res <- cbind(res,id=id)
        
        #remove rownames
        rownames(res) <- NULL
        
        #add column names
        colnames(res) <- c("nob","id")
        
        #re-arrange column names
        res = res[c("id", "nob")]
        
        #return the results dataframe
        res
        
}

complete("specdata",1)
complete("specdata",30:25)
complete("specdata",3)

```

## Part 3

Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. A prototype of this function follows

```{r assign1 p3 prototype}
corr <- function(directory, threshold = 0) {
        ## 'directory' is a character vector of length 1 indicating
        ## the location of the CSV files

        ## 'threshold' is a numeric vector of length 1 indicating the
        ## number of completely observed observations (on all
        ## variables) required to compute the correlation between
        ## nitrate and sulfate; the default is 0

        ## Return a numeric vector of correlations
}
```
For this function you will need to use the 'cor' function in R which calculates the correlation between two vectors. Please read the help page for this function via '?cor' and make sure that you know how to use it.

You can see some example output from this function. The function that you write should be able to match this output. Please save your code to a file named corr.R. To run the submit script for this part, make sure your working directory has the file corr.R in it. The output of this function should look like this:

```{r assign1 p3 output}

source("corr.R")
source("complete.R")
cr <- corr("specdata", 150)
head(cr)

## [1] -0.01896 -0.14051 -0.04390 -0.06816 -0.12351 -0.07589

summary(cr)

##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## -0.2110 -0.0500  0.0946  0.1250  0.2680  0.7630

cr <- corr("specdata", 400)
head(cr)

## [1] -0.01896 -0.04390 -0.06816 -0.07589  0.76313 -0.15783

summary(cr)

##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## -0.1760 -0.0311  0.1000  0.1400  0.2680  0.7630

cr <- corr("specdata", 5000)
summary(cr)

##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 

length(cr)

## [1] 0

cr <- corr("specdata")
summary(cr)

##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## -1.0000 -0.0528  0.1070  0.1370  0.2780  1.0000

length(cr)

## [1] 323

```

```{r assign p3 solution}

corr <- function(directory, threshold = 0) {
        
        #set the path
        path = "C:/Users/Fahad/Documents/R Projects/Coursera/R-Programming-Coursera/Data/"
        path = c(path,directory)
        path = paste(path,collapse="")
        
        #get the file List in that directory
        fileList = list.files(path)
        
        #import data
        Data = lapply(file.path(path,fileList),read.csv)
        
        #get the complete cases
        completeCases = lapply(Data, function(Data) Data[complete.cases(Data),])
        
        #create a dataframe to return the results
        res = numeric()
        
        
        #compare the no. of obs in complete cases against threshold and calculate 
        #correlations and save results in the res vector
        for(i in 1:length(completeCases)){
                if(nrow(completeCases[[i]])>=threshold){
                        res2 = data.frame()
                        res2 <- as.data.frame(completeCases[[i]])
                        res[i] <- cor(res2$sulfate, res2$nitrate)
                        rm(res2)
                
                }
                
        }
        res = res[!is.na(res)]
        res
}

```