Programming Assignment 1: Air Pollution: Instructions
========================================================

## Introduction

For this first programming assignment you will write three functions that are meant to interact with dataset that accompanies this assignment. The dataset is contained in a zip file **specdata.zip** that you can download from the Coursera web site. 

## Data

The zip file containing the data can be downloaded here: 

specdata.zip [2.4MB] 

The zip file contains 332 comma-separated-value (CSV) files containing pollution monitoring data for fine particulate matter (PM) air pollution at 332 locations in the United States. Each file contains data from a single monitor and the ID number for each monitor is contained in the file name. For example, data for monitor 200 is contained in the file "200.csv". Each file contains three variables:

* Date: the date of the observation in YYYY-MM-DD format (year-month-day)
* sulfate: the level of sulfate PM in the air on that date (measured in micrograms per cubic meter)
* nitrate: the level of nitrate PM in the air on that date (measured in micrograms per cubic meter)

For this programming assignment you will need to unzip this file and create the directory 'specdata'. Once you have unzipped the zip file, do not make any modifications to the files in the 'specdata' directory. In each file you'll notice that there are many days where either sulfate or nitrate (or both) are missing (coded as NA). This is common with air pollution monitoring data in the United States. 

## Part 1

Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA. A prototype of the function is as follows:

```{r prototype of Part 1}
pollutantmean <- function(directory, pollutant, id = 1:332) {
        ## 'directory' is a character vector of length 1 indicating
        ## the location of the CSV files
        
        ## 'pollutant' is a character vector of length 1 indicating
        ## the name of the pollutant for which we will calculate the
        ## mean; either "sulfate" or "nitrate".
        
        ## 'id' is an integer vector indicating the monitor ID numbers
        ## to be used
        
        ## Return the mean of the pollutant across all monitors list
        ## in the 'id' vector (ignoring NA values)
        }

```

The output should look like:

```{r part 1 output}

source("pollutantmean.R")
pollutantmean("specdata", "sulfate", 1:10)

## [1] 4.064

pollutantmean("specdata", "nitrate", 70:72)

## [1] 1.706

pollutantmean("specdata", "nitrate", 23)

## [1] 1.281


```

```{r simple function}

path = "C:/Users/Fahad/Documents/R Projects/Coursera/R-Programming-Coursera/Data/specdata"
fileList = list.files(path=path,pattern="*.csv",full.names=F)
all.data.files = lapply(fileList,read.csv,header=TRUE)

Data = sapply(fileList, function(file) {
                                 read.csv(file=paste(path, file, sep="/"), header=T)
                                 })
str(Data)
class(Data)
str(all.data.files)
class(all.data.files)


file.exists("C:/Users/Fahad/Documents/R Projects/Coursera/R-Programming-Coursera/Data/specdata/001.csv")
Examplecsv = read.csv("C:/Users/Fahad/Documents/R Projects/Coursera/R-Programming-Coursera/Data/specdata/023.csv",header=T)
str(Examplecsv)
summary(Examplecsv)
mean(Examplecsv$nitrate,na.rm=T)
paste(c("A","B"),1:10,sep="")

str(all.the.data)
DATA = do.call("rbind",all.files.data)
class(path)
id = 1
if(class(id)!="character"){
        id = as.character(id)
} else {
        paste(path,"/" + id + ".csv", collapse="/")
}

summary(DATA)
DATA$ID[1:10]
mean(DATA$sulfate[DATA$ID[1:10]],na.rm=T)
mean(DATA$sulfate[DATA$ID[DATA$ID==1:10]],na.rm=T)
DATA$sulfate[DATA$ID[DATA$ID==1:10]]
tapply(DATA$sulfate, DATA$ID,mean)
DATA$sulfate[DATA$ID==1:10]
DATA$ID[DATA$ID==1:10]
dim(DATA)
colMeans(DATA$sulfate[DATA$ID==23])
apply(DATA,2,mean)
```